Determination of Center of Rotation \\ in Computed Tomography

### 2017-07-15 seminar_cor/seminar_cor_haegele.tex
fan beam done, cone beam almnost

Unfortuantely, parallel beam geometry is impractical as a pencil beam \xray source (source emmiting only one ray in a specific direction) would have to be moved parallel to the detector to collect all samples for a single rotation angle, which is very time consuming.
However, parallel beam geometry occurs at synchrotron facilities \cite{something}.

\subsubsection{Fan Beam}
A more practical geometry is the fan beam, in which a source emits a flat beam that extends to the sides (like a fan blade). 
\Cref{subfig:fanbeam} shows a schematic of this geometry.
Here the projection for a single rotation angle can be obtained at once.

\subsubsection{Cone Beam}

>>>>

Unfortuantely, parallel beam geometry is impractical as a pencil beam \xray source (source emmiting only one ray in a specific direction) would have to be moved parallel to the detector to collect all samples for a single rotation angle, which is very time consuming \cite{Buzug2008_chap7}.
However, parallel beam geometry occurs at synchrotron facilities \cite{something}.

\subsubsection{Fan Beam}
A more practical geometry is the fan beam, in which a source emits a flat beam that extends to the sides (like a fan blade). 
\Cref{subfig:fanbeam} shows a schematic of this geometry.
Here the projection for a single rotation angle can be obtained at once. 
The geometry can be defined by the COR it rotates around, the offset of the source to the COR and its aperture angle.
In clinical practise curved detector rays are used, as shown in \cref{subfig:fanbeam}, to get constant aperture angle steps for each detector cell.
%
\begin{figure}[!h]
\centering
\subfloat[]{\includegraphics[width=0.43\linewidth]{eps/rebinning.eps}%
\label{subfig:rebinning1}}
\hfil
\subfloat[]{\includegraphics[width=0.51\linewidth]{eps/rebinning_geo.eps}%
\label{subfig:rebinning2}}
\caption{
On the left a schematic is shown with two fan beams with aperture angle $\varphi$ at different rotation angles.
It can be sen that the outer rays of the two are parallel.
The maximum distance of two parallel beams is limited by $\varphi$, and the maximum rotation angle $\theta$ at which two
parallel rays can be obtained is given by $\theta = \psi_{min}+\psi_{max}$ which are the angles from central to outer rays.
On the right a schematic is shown that points out the correspondence between fan and parallel beam geometry.
It can be seen that for the red ray which is the ray at rotation angle $\theta$ with deviation angle $\psi$ from the central ray, there exists an angle $\gamma$ in parallel beam geometry so that it is perpendicular to the $r$-axis (i.e. parallel beam detector).
Figures from~\cite{Buzug2008_chap7}
}
\label{fig:rebinning}
\end{figure}
%

As each ray projection for a single rotation angle in this geometry corresponds to a different ray directions (i.e. rays are not parallel here), the resulting sinogram differs from the sinogram of parallel beam geometry.
When analyzing this geometry further, it can be seen that the ray projections of a parallel beam geometry are inherent and can be found at different rotation angles.
Through a process called rebinning, the sinogram of the fan beam can be transformed to parallel beam.
\Cref{fig:rebinning} illustrates this concept.

\subsubsection{Cone Beam}
The previously introduced geometries are both 2-dimensional (2D) as they described projections within an x-y-plane onto a one dimensional detector.
To obtain a 3-dimensional (3D) image the scanned object would have to be moved perpendicular to the plane to obtain multiple slice images.
With the 3D cone beam geometry, this is not necessary as it uses a cone shaped beam that projects onto a 2D flat panel detector.
%
\begin{figure}
\centering
\includegraphics[width=\linewidth]{img/conebeam_centralslice.png}
\caption{
A cone beam setup with an object on a rotary table.
It can be seen that only the projection of the central slice is equivalent to a projection in fan beam geometry.
Other projections that are parallel to the s-axis of the detector use samples at multiple heights of the object.
Also the projection of the rotation axis onto the detector is shown which in this case is offset from detectors
z-axis.
Figure from \cite{Yang2011}
}
\label{fig:conebeam}
\end{figure}
%
In \cref{fig:conebeam} it can be seen that there exists a certain line on the detector which will receive the same input as a fan beam detector.
All other lines read projections through multiple planes of the object and are thus not rebinnable to parallel beam.
